EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-24 05_examples.tex
Merge updated from GitHub into OverLeaf. Co-authorhed-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\documentclass[edipack2.tex]{subfiles}
\begin{document}

@@

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
    \center{\bf Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   ...  
   !Read variables
   call ed_read_input('inputED.conf')
   !Solver-specific arays
   allocate(Smats(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Delta(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Hloc(Nspin,Nspin,Norb,Norb));Hloc=0d0
   ...
   !Initialize Bethe arrays
   allocate(Ebands(Le),Dbands(Le))
   Ebands = linspace(-D,D,Le,mesh=de)
   Dbands = dens_bethe(Ebands,Wband)*de
   !Set local H
   call ed_set_hloc(Hloc)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
    \center{\bf C++}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
#include <edipack_cbindings.h>
using namespace std;
...
//Main variables
int Le = 1000;
int iloop = 0;
double wmixing = 0.5;
double D = 1.0;
//Read  variables    
char input[] = "inputED.conf"; 
read_input(input);      
//Dimensions
int64_t d[4] = {Nspin,Nspin,Norb,Norb};
int total_size = d[0] * d[1] * d[2] * d[3];
int total_size_n5 = total_size * Lmats;    
// Solver-specific arrays
vector<complex<double>> Hloc(total_size);
vector<complex<double>> Smats(total_size_n5);
vector<complex<double>> Delta(total_size_n5);
...
// Initialize Bethe arrays
vector<double> Ebands, Dbands;
Ebands = linspace(-D,D,Le);
Dbands = dens_bethe(Ebands,Wband,de);
//set local H
ed_set_Hloc_single_N4(Hloc.data(), d);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

@@

\item[{\tiny $\blacksquare$}] Call the exact diagonalization {\bf impurity solver}
  whose only input is the set of parameters $\vec{x}$ contained in a
  rank-1 array. All  \NAME options are controlled through input file
  specifications.

\item[{\tiny $\blacksquare$}] Retrieve the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using dedicated function provided by \NAME API.

\item[{\tiny $\blacktriangleright$}] Evaluate the local interacting Green's function
  $G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e$ with
  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.

\item[{\tiny $\blacktriangleright$}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$.

  \item[{\tiny $\blacktriangleright$}] Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field,  potentially using the conjugate gradient procedures 
supplied by \NAME. Then, restart at step 1.
\end{itemize}

The first two steps, marked by the ${\tiny \blacksquare}$ symbol, 
are handled directly by \NAME routines. In the subsequent steps, 
indicated by the ${\tiny \blacktriangleright}$ symbol, control returns 
to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the 
critical importance of this optimization, \NAME provides access to 
well-tested functions, ensuring stability and reproducibility of 
results. Alternative optimization methods can also be employed as 
needed. 
An example of implementation is provided in the following listings.

@@

\paragraph{Results}
In the following, we present results for the interaction-driven Mott 
transition in the Bethe lattice, which captures the gradual 
transformation of a partially-filled metallic state into a correlated 
insulator.
To illustrate this, we report in panel (A) the evolution of the spectral 
function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of the 
interaction strength $U$. Despite the inherently {\it spiky} nature 
of the spectrum, resulting from the finite number of poles in the 
discretized effective bath, the characteristic features of the Mott 
transition are clearly visible. At low energies, a renormalized 
quasi-particle peak develops at the Fermi level ($\omega = 0$). 
Simultaneously, the system exhibits the formation of incoherent 
high-energy features, which eventually evolve into well-defined 
Hubbard bands in the Mott insulating phase for $U > U_\mathrm{c}$, 
with $U_\mathrm{c} \simeq 2.8D$.

@@

\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipack2py import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_\mathrm{loc}$ 
comm = MPI.COMM_WORLD
rank = comm.Get_rank()
master = (rank==0)
#Functions
def generate_kgrid(Nk):
    b1=2*np.pi*np.array([1.0,0.0])
    b2=2*np.pi*np.array([0.0,1.0])
    n1, n2 = np.meshgrid(np.arange(Nk), np.arange(Nk))
    n1=n1/Nk;n2=n2/Nk
    gridout = np.stack([n1.ravel(), n2.ravel()], axis=-1)
    return np.dot(gridout,[b1,b2])
def h_square2d(k,t):
  return -2*t*(np.cos(k[...,0,np.newaxis,np.newaxis])+np.cos(k[...,1,np.newaxis,np.newaxis]))*np.eye(ed.Norb)

@@

\paragraph{Results}
Here we showcase some results for the DMFT solution of the 
attractive Hubbard model across the BCS to BEC crossover regime, 
illustrating the capability of \NAME to handle $s$-wave 
superconductivity at both zero and finite temperatures.

@@

\subsection{Interacting Kane-Mele model (Fortran API, EDIpack2ineq layer)}

To complete the overview of \NAME capabilities, we consider another
model for interacting quantum spin-Hall insulators, defined on the honeycomb lattice, with two inequivalent sites in the unit-cell.

Neglecting the Rashba spin-orbit interaction and discarding any ionic character of the unit-cell, the noninteracting Kane-Mele Hamiltonian \cite{Kane2005PRLa,Kane2005PRL} can be written in momentum space as
\begin{align*}
    \mathbf{H}_\mathrm{KM} = \sum_{{k}} \psi^\dagger_{{k}} \mathbf{H}({k}) \psi_{{k}},
\end{align*}
with $\psi_{{k}} = [c_{{k},\mathrm{A},\up},\,c_{{k},\mathrm{B},\up},\,c_{{k},\mathrm{A},\dw},\,c_{{k},\mathrm{B},\dw}]$ and
    \begin{align}
    \mathbf{H}({k}) 
    &= x_{{k}}\Gamma_{01} + y_{{k}}\Gamma_{02} + \delta_{{k}}\Gamma_{33}, \label{eq:Hk_kanemele}\\[1mm]
   x_{{k}}&=t\bigl[\cos({k}\cdot{u})+\cos({k}\cdot{v})+1\bigr],\nonumber\\
   y_{{k}}&=t\bigl[\sin({k}\cdot{u})+\sin({k}\cdot{v})\bigr],\nonumber \\
   \delta_{{k}}&=2\lambda_\mathrm{so}\bigl[\sin({k}\cdot{u})-\sin({k}\cdot{v}) - \sin({k}\cdot{u} -{k}\cdot{v})\bigr], \nonumber
\end{align}
where ${u} = (3/2, \sqrt{3}/2)$, and ${v} = (3/2, -\sqrt{3}/2)$
are suitable basis vectors for the honeycomb lattice, $t$ is the
nearest-neighbor hopping amplitude and $\lambda_\mathrm{so}$ is 
the amplitude of a complex next-nearest-neighbor hopping amplitude,
with a spin-dependent $\pm\tfrac{\pi}{2}$ chiral phase arising from spin-orbit coupling and well-known to stabilize a topological gap at the Fermi level \cite{Kane2005PRLa,Kane2005PRL}.
The spinorial $4\times4$ matrices $\Gamma_{a\ell} = \sigma_a\otimes\tau_\ell$, are 
defined in terms of $\sigma_a$ and $\tau_\ell$ Pauli matrices, respectively 
referred to spin and sublattice degrees of freedom, with the convention that 
$\sigma_0$ and $\tau_0$ are $2\times2$ identity matrices.

@@

The essential features of the groundstate phase diagram have been established by intensive multi-method analyses \cite{Rachel2018ROPIP}.
For $\lambda_\mathrm{so}=0$ the model describes a Dirac semimetal, up to relatively high Hubbard repulsion, and magnetizes to an isotropic Néel state above above a critical $U/t$ ratio
\footnote{In contrast with the expectations for a generic half-filled bipartite lattice, for which the groundstate is antiferromagnetic for any nonzero interaction, this property descends from the peculiar properties of Dirac cones: the vanishing density of states
at the Fermi level renders ineffective the logarithmic instability, even in the presence of perfect nesting \cite{SorellaTosatti92,Graphene_RMP}.}.
With a finite spin-orbit coupling $\lambda_\mathrm{so}\neq0$, the quantum spin-Hall state results increasingly stable against magnetic ordering, due to its symmetry-protected nontrivial topology. Furthermore, the long-range ordering 
eventually stabilized at large $U/t$, is characterized by a lowered spin symmetry, due to the inherent coupling of spin and lattice degrees of freedom that favors in-plane easy-axis magnetization.
\cite{3dXY_KMH}, as predicted by Hartree-Fock and $t$-$J$ perturbative expansions \cite{Rachel2010PRBa,Rachel2010PRB} and
later confirmed by Auxiliary-Field Quantum Monte Carlo (AFQMC) 
simulations \cite{Hohenadler2013ROPIP}.

@@

The same off-diagonal spin symmetry must be realized also in
the bath Hamiltonian, which can be conveniently achieved with
either the \texttt{replica} and \texttt{general} choices for 
the \texttt{bath\_type} option. The most stable strategy to 
verify that the groundstate of the model is magnetized in the
plane also in DMFT is to allow either an $S_z$ or an $S_x$ (equivalently $S_y$) magnetization in the bath, and then 
compare the energies of the respective solutions at $T=0$. 
The corresponding Hamiltonian forms are:
\begin{align}
    \mathbf{H}^\mathrm{bath}_\perp &= 
        \sum_{p=1}^{N_\mathrm{b}}(\epsilon_p \sigma_0 + m_p \sigma_z), 
        \label{eq:afmz_dmft_ansatz} \\[1mm]
    \mathbf{H}^\mathrm{bath}_\parallel &= 
        \sum_{p=1}^{N_\mathrm{b}}(\epsilon_p \sigma_0 + m_p \sigma_x).
        \label{eq:afmx_dmft_ansatz}
\end{align}
comparing with \ref{eq:replica_bath_sym}, it is clear 
that this amount to having $N_\mathrm{sym}=2$, with
$\lambda_{1,p} \equiv \epsilon_p$ describing average
bath levels and $\lambda_{2,p} \equiv m_p$ allowing 
a magnetic split, respectively in the $S_z$ and $S_x$ spin bases.

@@

Hence, we can readily evaluate the total energy of both the AFM$_\perp$
and AFM$_\parallel$ calculations and compare them to establish which 
is favored as the groundstate of the system. In \figu{fig:KMenergy}(a)
we show results for the energy difference $E_\parallel-E_\perp$ across
a wide range of $U/t$ and $\lambda_\mathrm{so}$ values. To assess the
degree of accuracy of our total energy calculation, we benchmark against
Hartree-Fock (simple static mean-field, but inherently variational), 
DMFT with NRG solver and lattice AFQMC data, as taken from 
Ref.\,\cite{Racz2020}. The total energy comparison is reported in 
\figu{fig:KMenergy}(b), for $\lambda_\mathrm{so}=0$.

\begin{figure}
\hspace{1cm} (a) \hspace{7.25cm} (b)\\
    \includegraphics[width=0.47\linewidth,trim={0 0 0 5mm},clip]{figures/KMH_energy.pdf}\hfill
    \includegraphics[width=0.45\linewidth]{figures/benchEnergy_honey.pdf}\\
    \caption{In panel (a) the total energy difference between the in-plane 
    and off-plane AFM solutions found in DMFT-ED at $T=0$. Negative values
    signal a groundstate with in-plane magnetization. The plateau at zero
    energy difference marks the topological phase of the model (a 
    paramagnetic quantum spin-Hall insulator), at all points except on the $\lambda_\mathrm{so}=0$ line, where the groundstate is an interacting
    Dirac liquid, at weak coupling, or an isotropic antiferromagnet at strong coupling. In panel (b) we compare the total energies across the 
    $\lambda_\mathrm{so}=0$ line to published data for Hartree-Fock,
    DMFT-NRG and AFQMC solutions \cite{Racz2020}.}
    \label{fig:KMenergy}
\end{figure}

>>>>

\documentclass[edipack_sp.tex]{subfiles}
\begin{document}

@@

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   !...  
   !Read variables
   call ed_read_input('inputED.conf')
   !Solver-specific arays
   allocate(Smats(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Delta(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Hloc(Nspin,Nspin,Norb,Norb))
   Hloc=0d0
   !...
   !Initialize Bethe arrays
   allocate(Ebands(Le),Dbands(Le))
   Ebands = linspace(-D,D,Le,mesh=de)
   Dbands = dens_bethe(Ebands,Wband)*de
   !Set local H
   call ed_set_hloc(Hloc)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
#include <edipack_cbindings.h>
using namespace std;
!...
//Main variables
int Le = 1000;
int iloop = 0;
double wmixing = 0.5;
double D = 1.0;
//Read  variables    
char input[] = "inputED.conf"; 
read_input(input);      
//Dimensions
int64_t d[4] = {Nspin,Nspin,Norb,Norb};
int total_size = d[0] * d[1] * d[2] * d[3];
int total_size_n5 = total_size * Lmats;    
// Solver-specific arrays
vector<complex<double>> Hloc(total_size);
vector<complex<double>> Smats(total_size_n5);
vector<complex<double>> Delta(total_size_n5);
...
// Initialize Bethe arrays
vector<double> Ebands, Dbands;
Ebands = linspace(-D,D,Le);
Dbands = dens_bethe(Ebands,Wband,de);
//set local H
ed_set_Hloc_single_N4(Hloc.data(), d);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list2}
{\bf Fortran} & {\bf C++}\\
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

@@

\item[{\tiny {\bf EDIpack}}] Call the exact diagonalization {\bf impurity solver}
  whose only input is the set of parameters $\vec{x}$ contained in a
  rank-1 array. All  \NAME options are controlled through input file
  specifications.

\item[{\tiny {\bf EDIpack}}] Retrieve the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using dedicated function provided by \NAME API.

\item[{\tiny {\it User}}] Evaluate the local interacting Green's function
  $G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e$ with
  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.

\item[{\tiny {\it User}}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$.

  \item[{\tiny ${\it User} $>$ {\bf EDIpack}}] Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field,  potentially using the conjugate gradient procedures 
supplied by \NAME. Then, restart at step 1.
\end{itemize}

The first two steps are handled directly by \NAME routines, while during the subsequent steps the 
 control returns to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the 
critical importance of this optimization, \NAME provides access to 
well-tested functions, ensuring stability and reproducibility of 
results. Alternative optimization methods can also be employed as 
needed. 
An example of implementation is provided in the following listings.

@@

\paragraph{Results.}
In the following, we present results for the interaction-driven Mott 
transition in the Bethe lattice, which captures the gradual 
transformation of a partially-filled metallic state into a correlated 
insulator.
To illustrate this, we report in panel (A) the evolution of the spectral 
function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of the 
interaction strength $U$. Despite the inherently {\it spiky} nature 
of the spectrum, resulting from the finite number of poles in the 
discretized effective bath, the characteristic features of the Mott 
transition are clearly visible. At low energies, a renormalized 
quasi-particle peak develops at the Fermi level ($\omega = 0$). 
Simultaneously, the system exhibits the formation of incoherent 
high-energy features, which eventually evolve into well-defined 
Hubbard bands in the Mott insulating phase for $U > U_\mathrm{c}$, 
with $U_\mathrm{c} \simeq 2.8D$.

@@

\begin{lstlisting}[style=mypython,numbers=none,basicstyle={\scriptsize\ttfamily}]
import numpy as np
from edipack2py import global_env as ed
import mpi4py
from mpi4py import MPI
import os,sys
from aux_funx import * #include function to build $G_\mathrm{loc}$ 
comm = MPI.COMM_WORLD
rank = comm.Get_rank()
master = (rank==0)
#Functions
def generate_kgrid(Nk):
    b1=2*np.pi*np.array([1.0,0.0])
    b2=2*np.pi*np.array([0.0,1.0])
    n1, n2 = np.meshgrid(np.arange(Nk), np.arange(Nk))
    n1=n1/Nk;n2=n2/Nk
    gridout = np.stack([n1.ravel(), n2.ravel()], axis=-1)
    return np.dot(gridout,[b1,b2])
def h_square2d(k,t):
  return -2*t*( np.cos(k[...,0,np.newaxis,np.newaxis])+
                np.cos(k[...,1,np.newaxis,np.newaxis]))*np.eye(ed.Norb)

@@

\paragraph{Results.}
Here we showcase some results for the DMFT solution of the 
attractive Hubbard model across the BCS to BEC crossover regime, 
illustrating the capability of \NAME to handle $s$-wave 
superconductivity at both zero and finite temperatures.

@@

\subsection{Interacting Kane-Mele model (Fortran API, EDIpack2ineq layer, iRDM)}

To complete the overview of \NAME capabilities, we consider another
model for interacting quantum spin-Hall insulators, defined on the honeycomb lattice, with two inequivalent sites in the unit-cell. Neglecting the Rashba spin-orbit interaction and discarding any ionic character of the unit-cell, the noninteracting Kane-Mele Hamiltonian \cite{Kane2005PRLa,Kane2005PRL} can be written in momentum space as
\begin{align*}
    \mathbf{H}_\mathrm{KM} = \sum_{{k}} \psi^\dagger_{{k}} \mathbf{H}({k}) \psi_{{k}},
\end{align*}
with $\psi_{{k}} = [c_{{k},\mathrm{A},\up},\,c_{{k},\mathrm{B},\up},\,c_{{k},\mathrm{A},\dw},\,c_{{k},\mathrm{B},\dw}]$ and
    \begin{align}
    \mathbf{H}({k}) 
    &= x_{{k}}\Gamma_{01} + y_{{k}}\Gamma_{02} + \delta_{{k}}\Gamma_{33}, \label{eq:Hk_kanemele}\\[1mm]
   x_{{k}}&=t\bigl[\cos({k}\cdot{u})+\cos({k}\cdot{v})+1\bigr],\nonumber\\
   y_{{k}}&=t\bigl[\sin({k}\cdot{u})+\sin({k}\cdot{v})\bigr],\nonumber \\
   \delta_{{k}}&=2\lambda_\mathrm{so}\bigl[\sin({k}\cdot{u})-\sin({k}\cdot{v}) - \sin({k}\cdot{u} -{k}\cdot{v})\bigr], \nonumber
\end{align}
where ${u} = (3/2, \sqrt{3}/2)$, and ${v} = (3/2, -\sqrt{3}/2)$
are suitable basis vectors for the honeycomb lattice, $t$ is the
nearest-neighbor hopping amplitude and $\lambda_\mathrm{so}$ is 
the amplitude of a complex next-nearest-neighbor hopping,
with a spin-dependent $\pm\tfrac{\pi}{2}$ chiral phase arising from spin-orbit coupling, which is well-known to open a topological gap at the Fermi level \cite{Kane2005PRLa,Kane2005PRL}.
The spinorial $4\times4$ matrices $\Gamma_{a\ell} = \sigma_a\otimes\tau_\ell$, are 
defined in terms of $\sigma_a$ and $\tau_\ell$ Pauli matrices, respectively 
referred to spin and sublattice degrees of freedom, with the convention that 
$\sigma_0$ and $\tau_0$ are $2\times2$ identity matrices.

@@

The essential features of the groundstate phase diagram have been established by intensive multi-method analyses \cite{Rachel2018ROPIP}.
For $\lambda_\mathrm{so}=0$ the model describes a Dirac semimetal, up to relatively high Hubbard repulsion, and magnetizes to an isotropic Néel state above above a critical $U/t$ ratio\footnote{In contrast with the expectations for a generic half-filled bipartite lattice, for which the groundstate is antiferromagnetic for any nonzero interaction, this property descends from the peculiar properties of Dirac cones: the vanishing density of states
at the Fermi level renders ineffective the logarithmic instability, even in the presence of perfect nesting \cite{SorellaTosatti92,Graphene_RMP}.}.
With a finite spin-orbit coupling $\lambda_\mathrm{so}\neq0$, the quantum spin-Hall state results increasingly stable against magnetic ordering, due to its symmetry-protected nontrivial topology. Furthermore, the long-range ordering 
eventually stabilized at large $U/t$, is characterized by a lowered spin symmetry, due to the inherent coupling of spin and lattice degrees of freedom that favors in-plane easy-axis magnetization
\cite{3dXY_KMH}, as predicted by Hartree-Fock and $t$-$J$ perturbative expansions \cite{Rachel2010PRBa,Rachel2010PRB} and
later confirmed by Auxiliary-Field Quantum Monte Carlo (AFQMC) 
simulations \cite{Hohenadler2013ROPIP}.

@@

The same off-diagonal spin symmetry must be realized also in
the bath Hamiltonian, which can be conveniently achieved with
either the \texttt{replica} and \texttt{general} choices for 
the \texttt{bath\_type} option. The most stable strategy to 
verify that the groundstate of the model is magnetized in the
plane also in DMFT is to allow either an $S_z$ or an $S_x$ (equivalently $S_y$) magnetization in the bath, and then 
compare the energies of the respective solutions at $T=0$. 
The corresponding Hamiltonian forms are:
\begin{align}
    \mathbf{H}^\mathrm{bath}_\perp &= 
        \sum_{p=1}^{N_\mathrm{b}}(\epsilon_p \sigma_0 + m_p \sigma_z), 
        \label{eq:afmz_dmft_ansatz} \\[1mm]
    \mathbf{H}^\mathrm{bath}_\parallel &= 
        \sum_{p=1}^{N_\mathrm{b}}(\epsilon_p \sigma_0 + m_p \sigma_x).
        \label{eq:afmx_dmft_ansatz}
\end{align}
comparing with \ref{eq:replica_bath_sym}, it is clear 
that this amounts to having $N_\mathrm{sym}=2$, with
$\lambda_{1,p} \equiv \epsilon_p$ as the degenerate
bath levels that are Zeeman split by $\lambda_{2,p} \equiv m_p$, respectively in the $S_z$ and $S_x$ bases.

@@

Hence, we can readily evaluate the total energy of both the AFM$_\perp$
and AFM$_\parallel$ calculations and compare them to establish which 
is favored as the groundstate of the system. In \figu{fig:KMenergy}(a)
we show results for the energy difference $E_\parallel-E_\perp$ across
a wide range of $U/t$ and $\lambda_\mathrm{so}$ values. To assess the
degree of accuracy of our total energy calculation, we report in
\figu{fig:KMenergy}(b) a $\lambda_\mathrm{so}=0$ benchmark against
Hartree-Fock (simple static mean-field, but inherently variational), 
DMFT with NRG solver and lattice AFQMC data, as taken from 
Ref.\,\cite{Racz2020}.

\begin{figure}
\hspace{1cm} (a) \hspace{7.25cm} (b)\\
    \includegraphics[width=0.47\linewidth,trim={0 0 0 5mm},clip]{figures/KMH_energy.pdf}\hfill
    \includegraphics[width=0.45\linewidth]{figures/benchEnergy_honey.pdf}\\
    \caption{In panel (a) we report the total energy difference between 
    the in-plane and the
    off-plane AFM solutions found with \NAME at $T=0$. Negative values
    signal a groundstate with in-plane magnetization. The plateau at zero
    energy difference marks the topological phase of the model (a 
    paramagnetic quantum spin-Hall insulator), at all points except on the $\lambda_\mathrm{so}=0$ line, where the groundstate is an interacting
    Dirac liquid, at weak coupling, or an isotropic antiferromagnet at strong coupling. In panel (b) we compare the total energies, across the 
    $\lambda_\mathrm{so}=0$ line, to published data for Hartree-Fock,
    DMFT-NRG and AFQMC solutions \cite{Racz2020}.}
    \label{fig:KMenergy}
\end{figure}

\begin{figure}
    \centering
    (a) \hspace{12cm} (b) \\
    \includegraphics[width=0.8\linewidth]{figures/flakes_edipack.pdf}\\[3mm]
    (c) \hspace{12cm} (d) \\
    \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU4.5_x.pdf} \hspace{1.2cm}
    \includegraphics[width=0.33\linewidth]{figures/3NflakeU6_x.pdf} \vspace{3mm}
    \caption{Kane-Mele-Hubbard model at the nanoscale.
    In (a) the local in-plane
    magnetization $n_{i,\rightarrow}-n_{i,\leftarrow}$. In
    (b) the local correlation as measured by the mutual 
    information between local in-plane spin eigenstates
    $I(\rightarrow:\leftarrow)$ \cite{BellomiaPhD}. Both the
    (a)-(b) panels mark the edge and "bulk" sites in lighter 
    and darker colors, respectively.
    In (c) and (d) a visual representation of the given 
    nano-system at different values of $U/t$, as marked 
    respectively by the orange and red vertical lines 
    in panels (a) and (b). The area of the 
    circles is proportional to $I(\rightarrow:\leftarrow)$, 
    the length of the arrows to the local magnetization.
    All data are taken at $\lambda_\mathrm{so}=0.3t$ and $T=0$.}
    \label{fig:KMflake}
\end{figure}

\subsubsection{Kane-Mele Nanoflakes}
The capabilities of the EDIpack2ineq layer go well beyond the
simple case of a two-site unit-cell. We can readily adapt the
code to deal with nanoscopic systems such as hexagonal flakes
\cite{Valli2016PRB,Valli2018NL}, by building a
$N_\mathrm{ineq}\times N_\mathrm{ineq}$ local Hamiltonian 
$H_\mathrm{loc}$ with open boundary conditions, and implementing
the self-consistent equations with a trivial single momentum 
$k=0$. With respect to the infinite lattice, it is expected that
the system develops an inhomogeneous magnetization, so the number
of impurity models to solve cannot be halved by imposing the usual
Néel symmetry. Some lowered point symmetry can be found, but a 
safer approach of solving all the $N_\mathrm{ineq}$ impurity models
leads to good results, with fair control of spatial numerical noise.

In \figu{fig:KMflake} we report results for a nanosystem of 
$N_\mathrm{ineq}=54$ sites, solved with the bath parametrization 
described in \equ{eq:afmx_dmft_ansatz}, for the in-plane 
antiferromagnetic state.
Panel (a) shows the interaction dependence of the local magnetization,
computed from the impurity groundstate as  
$ \ibra c^\dagger_{i,\up} c_{i,\dw} + c^\dagger_{i,\dw} c_{i,\up} \iket$,
so that we can define the $S_x$ spin-occupation numbers as
\begin{equation*}
    n_{i,\rightarrow} \equiv \frac{n_{i,\up} + n_{i,\dw} +
    \ibra c^\dagger_{i,\up} c_{i,\dw} + c^\dagger_{i,\dw} c_{i,\up} \iket}{2},
    \qquad
    n_{i,\leftarrow} \equiv \frac{n_{i,\up} + n_{i,\dw} -
    \ibra c^\dagger_{i,\up} c_{i,\dw} + c^\dagger_{i,\dw} c_{i,\up} \iket}{2},
\end{equation*}
which trivially satisfy the conservation of the local average charge $n_{i,\rightarrow} + n_{i,\leftarrow} = n_{i,\up} + n_{i,\dw}$.
Having a direct expression for $n_{i,\rightarrow}$ and $n_{i,\leftarrow}$
we can verify that the impurity reduced density matrix, computed by 
tracing the bath states in the \texttt{ed\_mode=nonsu2} representation 
(see \ref{sSecRDM}), once diagonalized, takes the familiar form
\cite{XiDai2013,WalshLetter2019,BellomiaPhD,BellomiaKMH,Bellomia_intracorr} 
\begin{equation}
    \rho^\mathrm{imp} = 
     \begin{pmatrix}
            \left\ibra\left(1-n_{i,\rightarrow}\right)\left(1-n_{i,\leftarrow}\right)\right\iket   &    0          &       0       &   0 \\
                    0       &    \left\ibra n_{i,\rightarrow}\left(1-n_{i,\leftarrow}\right)\right\iket  &       0       &   0 \\
                   0       &       0   &    \left\ibra\left(1-n_{i,\rightarrow}\right)n_{i,\leftarrow}\right\iket      &   0 \\
                    0       &       0   &       0          &    \left\ibra n_{i,\rightarrow} n_{i,\leftarrow}\right\iket
        \end{pmatrix} 
\end{equation}
in the basis of the \emph{natural} spin-orbitals, whose occupation numbers
are given by $n_{i,\rightarrow}$ and $n_{i,\leftarrow}$.

As extensively discussed in \cite{BellomiaPhD,BellomiaKMH,Bellomia_intracorr},
by defining the spin traces of $\rho^\mathrm{imp}$, as
\begin{equation}
    \rho^\mathrm{imp}_\sigma = 
    \Tr_{\bar{\sigma}} \left[\rho^\mathrm{imp}\right] =
    \begin{pmatrix}
        \ibra n_{i,\sigma} \iket & 0 \\
        0 & \ibra n_{i,\bar{\sigma}} \iket
    \end{pmatrix}
\end{equation}
we can directly measure local correlations by defining the mutual information
between the impurity spin-orbitals
$I(\rightarrow:\leftarrow) = 
    S(\rho^\mathrm{imp}_{\rightarrow}) + 
    S(\rho^\mathrm{imp}_{\leftarrow}) -
    S(\rho^\mathrm{imp})$,
where $S(\cdot)$ denotes the von Neumann entropy of its argument.

In panel (b) of \figu{fig:KMflake} we report the spatially resolved
interaction dependency of $I(\rightarrow:\leftarrow)$, as a direct 
observation of the enhanced local correlations at the boundary of the
finite system. While it is usually assumed that edge sites must be
more correlated than internal ("bulk") sites, based on the reduced
lattice coordination, the inter-spin mutual information beautifully
highlights in a quantitative fashion the phenomenon: the edge correlations,
marked in lighter green, are decidedly higher as long as the system 
remains in a weakly magnetized regime. When the local magnetization
exceeds half of its saturation value, all local correlations start to decrease,
as spin fluctuations are damped by ordering. Significantly, the edge 
correlations get lower than the bulk correlations as the maximum is 
crossed, consistently with the observation that their magnetization
is always higher and so their fluctuations are frozen faster.
Finally, we observe that the nanosystem of \figu{fig:KMflake} has 
finite local magnetization for arbitrary low interaction strengths,
in striking contrast with the infinite lattice. This is a well-known
effect of quantum confinement, as proposed in early works on graphene
flakes \cite{Valli2016PRB,Valli2018NL}.
